### 2011-05-05 paper/paper.tex
Added adaptive methods section text

~~~ We consider the problem of randomly and uniformly sampling a point from a convex region in $n$-dimensional space. This has applications, for instance, in convex optimization \cite{Dabbene}. We phrase the problem as follows: we are given a convex region $\mathcal{R}$ in the form of a \emph{membership oracle}, a black box which can determine whether any point $x$ is in $\mathcal{R}$, along with a starting point $x_0$ in $\mathcal{R}$. The task is to sample a point from the region with with a distribution as close to uniform as possible.

@@

>>>>

We consider the problem of randomly and uniformly sampling a point from a convex region in $n$-dimensional space. This has applications, for instance, in convex optimization \cite{Dabbene}. We phrase the problem as follows: we are given a convex region $\mathcal{R}$ in the form of a \emph{membership oracle}, a black box which can determine whether any point $x$ is in $\mathcal{R}$, along with a starting point $x_0$ in $\mathcal{R}$. The task is to sample a point from the region with with a distribution as close to uniform as possible.

@@

Metropolis Markov chain methods, which include the basic ball walking method, have two major factors which can cause increased runtime. The first and more theoretically important is that if the current point is close to a corner of the region, then the chance of a step succeeding can be exponentially small. This arises even in simple cases such as a cube or a simplex if the sampling algorithm is given a point near a corner as the initial point. It also can occur if the step size is too large, the extreme case of which is sampling from a ball that is much larger than the entire admissible region.

In this case shrinking the step size would clearly help, but it is not difficult to see that a small step size also mitigates the effect of nearby corners, as a larger fraction of the surrounding ball lies inside the admissible region.

The other contribution to the length of time required for Metropolis methods is that if the step size is too small, it takes a long time to walk from one point in the admissible region to another. As it is certainly not possible for the Markov chain to mix before it is able to reach any other given point, a small step size causes the state space to have a large diameter, and thus a long mixing time.

In order to find a good balance between the two main contributors to runtime stated above, we introduce an adaptive walk, which we run for some time before running the standard ball walking or Metropolis method. This adaptive walk uses the success or failure of a step to guess whether its current step size is too small or too large, and then compensates by modifying it by a constant factor.

In the case of ball walking, our step size is the radius of the ball around the current point from which we draw the next point, and in the Metropolis method the step size corresponds to the standard deviation of the normal distributions used. However, it is not clear that this procedure has the uniform distribution as its stationary distribution, so we follow this adaptive procedure with the nonadaptive version, using the step size left by the adaptive method.

Since the nonadaptive methods converge to the stationary distribution from any starting distribution, this yields a method that converges to the stationary distribution, but also attempts to find a good step size in order to converge more rapidly.
